### 2023-06-07 Conference Submission/DraftDoc.tex
finished draft

\begin{section}{Introduction}
    The modelling of temporal networks is an important task in many real world applications including symptom interactions for mental health \cite{jordan2020current,contreras2020temporal}, epidemiology \cite{masuda2013predicting}, and protein interactions \cite{lucas2021inferring,jin2009identifying}.
    Temporal networks can be seen as dynamical systems: that is a system in which we have points, in our case nodes in a network, whose states, the edges connecting them, that vary dependent in time.
    Discovering the underlying equations governing these dynamical systems proves challenging, because changes in network structure are typically observed in the form of discrete jumps from one state to another, for example an edge between two nodes not being observed at the first time step then appearing at the next.
    Here, we propose a hybrid statistical and deep learning framework that allows us to model temporal networks as continuous-time dynamical systems, discover a fitting set of differential equations describing it, and, exploiting that discovery, predict the time evolution of a network.

    Differential equations are useful for modelling systems where the state of one variable can effect the trajectories of other variables. We observe this behavior in temporal networks; nodes' connections within the network can influence the formation and decay of edges between other nodes, for example the phenomenon of preferential attachment observed in \cite{newman2001clustering,capocci2006preferential}, where a node is more lively to gain the connections the more connections it has. With this in mind we might wish to draw on the rich mathematical literature of differential equation modelling.

    In the common representation of networks as binary-valued adjacency matrices, the events recorded in a temporal sequence of networks correspond to topological events, such as the appearance or the disappearance of link.
    Because of the discrete nature of events, directly modelling the temporal networks as dynamical systems would require us to handle discrete jumps.
    The topological nature of temporal networks, and the discontinuous character of their temporal evolution, make it challenging to use differential equations techniques.

    Here, we overcome the discreteness problem by interpreting networks as Random Dot Product Graphs; a well established statistical model for complex networks, that embeds nodes in a continuous, low-dimensional metric space by using a truncated singular value decomposition\cite{athreya2017statistical}. In this way we translate the hard problem of modelling discrete events in the space of networks to the easier problem of modelling continuous change in the embedding space. We then define and use systems of Neural Network Differential Equations (NNDE)\cite{SciML_C_Rak} to approximate the time evolution of the embedding space, and symbolic regression techniques to discover the functional form of the fitted NNDEs. These functional forms are interpretable (as they read as classic differential equations) and allow us to predict forward in time the evolution of the temporal networks.

    In this manuscript, we prove that the temporal network prediction problem can be successfully re-interpreted as a dynamical system modelling problem by taking the singular value decomposition of a sequence of adjacency matrices and training a NNDE to model an approximation to the underlying differential equation. We then go on to create a symbolic equation of this approximation.

    We apply our proposed framework to three small example temporal networks with the hope of exploring the limitations and strengths of the proposed framework. 
    The framework we are introducing is extremely flexible, and our research regarding the optimal structure of the Neural Networks used for the NNDEs is just started.
    We are confident that future research can identify more fitting Neural Network structures than the simple one adopted here.
    For this reason, we did not yet attempt to benchmark our model against other classic temporal network prediction methods.
    As it is completely general, we believe that the framework we are introducing can be usefully applied to areas of medicine, especially protein interaction networks; population dynamics for network ecology; and social network modelling. In particular, we discuss how specific domain knowledge relative to the prediction scenario can be taken into account, moving from NNDEs to Universal Differential Equations.
\end{section}

\begin{section}{Methodology}

        We are given a sequence of graphs, and our goal is to predict the next graph in the sequence. To achieve this, we reinterpret each graph as a sequence of points in a latent space; that is, a space in which similar vertices (ones with many of the same neighbours) are mapped to a similar point in space. We then Train a Neural Network Differential Equation (NNDE) to approximate the rate of change of the points in this latent space. With this numerical equation we can then use symbolic regression techniques to find a functional equation that matches the numerical equation. Using this process we discover an interpretable function that governs the evolution of a differential equation.

        \begin{figure}
            \centering
            \begin{subfigure}[c]{1\textwidth}
                \begin{tabular}{llll}
                \begin{subfigure}[c]{0.25\textwidth}
                    \centering
                    \resizebox{.6\width}{!}{\input{SynthDataDiagrams/methodology1.tex}}
                    \label{method net, a}
                \end{subfigure}
                &
                \centering
                \begin{subfigure}[c]{0.25\textwidth}
                    \centering
                    \resizebox{.6\width}{!}{\input{SynthDataDiagrams/methodology2.tex}}
                    \label{method net, b}
                \end{subfigure}
                &
                $\cdots$
                &
                \centering
                \begin{subfigure}[c]{0.25\textwidth}
                    \centering
                    \resizebox{.6\width}{!}{\input{SynthDataDiagrams/methodology3.tex}}
                    \label{method net, c}
                \end{subfigure}

                \end{tabular}
                \caption{Example sequence of networks.}
                \label{method nets}

            \end{subfigure}
            \begin{subfigure}[c]{1\textwidth}
                \begin{tabular}{llll}
                    $\begin{bmatrix}
                        \cdot & 1 & 1 & 1 & \cdot & \cdot & \cdot & \cdot\\
                        1 & \cdot & 1 & \cdot & \cdot & \cdot & \cdot & \cdot\\
                        1 & 1 & \cdot & 1 & 1 & \cdot & \cdot & \cdot\\
                        1 & \cdot & 1 & \cdot & \cdot & \cdot & \cdot & \cdot\\
                        \cdot & \cdot & 1 & \cdot & \cdot & 1 & 1 & \cdot\\
                        \cdot & \cdot & \cdot & \cdot & 1 & \cdot & \cdot & \cdot\\
                        \cdot & \cdot & \cdot & \cdot & 1 & \cdot & \cdot & 1\\
                        \cdot & \cdot & \cdot & \cdot & \cdot & \cdot & 1 & \cdot
                    \end{bmatrix}$
                    &
                    $\begin{bmatrix}
                        \cdot & 1 & 1 & 1 & \cdot & \cdot & \cdot & \cdot\\
                        1 & \cdot & 1 & 1 & \cdot & \cdot & \cdot & \cdot\\
                        1 & 1 & \cdot & 1 & 1 & \cdot & \cdot & \cdot\\
                        1 & 1 & 1 & \cdot & \cdot & \cdot & \cdot & \cdot\\
                        \cdot & \cdot & 1 & \cdot & \cdot & 1 & 1 & \cdot\\
                        \cdot & \cdot & \cdot & \cdot & 1 & \cdot & \cdot & 1\\
                        \cdot & \cdot & \cdot & \cdot & 1 & \cdot & \cdot & 1\\
                        \cdot & \cdot & \cdot & \cdot & \cdot & 1 & 1 & \cdot
                    \end{bmatrix}$
                    &
                    $\cdots$
                    &
                    $\begin{bmatrix}
                        \cdot & 1 & 1 & 1 & \cdot & \cdot & \cdot & \cdot\\
                        1 & \cdot & 1 & 1 & \cdot & \cdot & \cdot & \cdot\\
                        1 & 1 & \cdot & 1 & \cdot & \cdot & \cdot & \cdot\\
                        1 & 1 & 1 & \cdot & \cdot & \cdot & \cdot & \cdot\\
                        \cdot & \cdot & \cdot & \cdot & \cdot & 1 & 1 & \cdot\\
                        \cdot & \cdot & \cdot & \cdot & 1 & \cdot & 1 & 1\\
                        \cdot & \cdot & \cdot & \cdot & 1 & 1 & \cdot & 1\\
                        \cdot & \cdot & \cdot & \cdot & \cdot & 1 & 1 & \cdot
                    \end{bmatrix}$

                \end{tabular}
                \caption{Sequence of sparse adjacency matrices associated with the networks in \autoref{method nets}.}
                \label{method adjacency}
            \end{subfigure}

            \begin{tabular}{llll}

            \end{tabular}

            \caption{Visual illustration of the proposed framework.}
            \label{framework illustration}
        \end{figure}

        \subsection{Singular Value Decomposition and Random Dot Product Graphs}
            \label{svd}
            In this section we discuss the Singular Value Decomposition (SVD) of a matrix, and its usefulness in creating Random Dot Product Graphs (RDPG).

            We define the matrix $A$ to be a real $m \times n$ with $n \le m$. $A$ can be expressed as \cite{forsythe1967computer}:
            \begin{align}
                A&=L\Sigma R'
            \end{align}
            Where $L, R$ are real valued, orthonormal matrices. The columns of $L$ consist of the $n$ largest eigenvectors of $AA'$ and the columns $R$ consist of the eigenvectors of $A'A$. $\Sigma$ is a diagonal matrix whose entries are the square root of the positive eigenvalues of $A'A$ in decreasing order.

            In order to reduce the problem size, we can truncate the $L,R,\Sigma$ to the first $d$ columns to yield, $\hat L, \hat R, \hat \Sigma$ such that:

            \begin{align}
                A &\approx \hat L \hat \Sigma \hat R' \\
                \hat A &= \hat L \hat \Sigma \hat R'
                \label{approx}
            \end{align}

            The approximation of $A$ is defined as $\hat A$ and is given by \autoref{approx} can be interpreted as a graph where the entry $(i,j)$ of the matrix is the probability that an edge exists from $i \rightarrow j$. This approximation of $A$ is known as an RDPG.

            It is important to note that the SVD is not unique. The distances between points will be the same for every embedding, but the orientation may be different, and so, if the SVD of each time step is not aligned, learning will be effectively impossible as we will essentially be trying to model random noise. Here, we can use Procrustes alignment XXX cite package XXX to align the matrices after they have been generated.

            The SVD gives is an RDPG\cite{athreya2017statistical}. The SVD can be considered to give a latent space representation of the matrix\cite{hoff2002latent}. That is, the rows of the matrices $L,R$ can be viewed as vector points in a $d$-dimensional space, where $d$ is the number of singular values we take for our SVD truncation. In this space similar nodes in the network are mapped to similar points. A visual representation of this process can be seen in \autoref{framework illustration}.

            When obtaining the RDPGs for our temporal network, we take each network observation \autoref{method nets} and represent them in the form of an adjacency matrix \autoref{method adjacency}. We then use a truncated singular value decomposition to reduce the dimensionality of our adjacency matrices\cite{golub1971singular}. This truncation introduces some loss, but can dramatically reduce the size of our problem. \Citeauthor{Runghen2021}\cite{Runghen2021} effectively employ this method to reduce a problem involving over 130,000 visitors to a 6 dimensional latent space, whilst preserving 70\% of the information.

            To execute the SVD we use the ARPACK Julia package \cite{lehoucq1998arpack}. This package uses an iterative approach to approximate the eigenvectors and singular values of each of our adjacency matrices\cite{lehoucq1996deflation}.

    \subsection{Neural Network Differential Equation}
        Once we have the SVD embedding of points, we look to model the trajectories of a target point. That is, with some input data, point location, distance to neighbours, etc, we want to model the differential equation that governs the movement of a point. Notably this method can be used to model multiple points by simply changing the node being predicted and trained on.

        Our problem can be defined as:
        \begin{align}
            A_{t+1} = f(A(t),t)
        \end{align}
        Where $A_t$ is the adjacency matrix of the temporal network at time $t$. By using the SVD to obtain the approximation $\hat A_t$, which can evolve continuously, we can reinterpret our problem as:

        \begin{align}
            \frac{d\hat A}{dt} &= g(\hat A(t), t)
        \end{align}

        In this thesis, as a proof of concept, we focus on a modelling the change in structure of a single node, referred to as the target node $u$. In this case, the form of the problem becomes:

        \begin{align}
            \frac{du}{dt} &= g(u(t), t)
        \end{align}

        Because we are only modelling the movement of a single node, we assume that we have access to the rest of the network structure at every time step for the prediction. This would likely not be the case in a real world application, but will allow us to observe some of the strengths and weaknesses of this framework in a controlled setting. If we wanted to predict the whole network we would extrapolate out every point and use the dot product of each point to predict the edges.

        Here we train a neural network to approximate the function $g(u(t),t)=NN(u(t),t,\theta)$. Where $\theta$ is the parameters of the neural network.

        To obtain this differential equation we make use of the SciML ecosystem of packages in the Julia programming language \cite{SciML_C_Rak}.

        % KNN Explaination
        As a way of limiting the complexity of our model, we limit the input of the neural network to the euclidean distances between the target node and its $k$ nearest neighbours. The target node is the node that we are looking to model the movement of and stays the same throughout each of the temporal networks. 
        % \begin{align}
    %     %\centering
    %     \mathbb{P}(i\rightarrow j)&\coloneqq L_i \cdot R_j\\
    %     \mathbb{P}(\hat{i}\rightarrow j)&\coloneqq L_{\hat{i}} \cdot R_j
    % \end{align}
    % Where $(i,j)$ are nodes in the network and $\hat{i}$ is another node at the same timestep to $i$. This implies that the distance between these two nodes with respect to node $j$ is given by
    % \begin{equation}
    %     \begin{gathered}
    %         %\centering
    %         ||\mathbb{P}(i\rightarrow j)-\mathbb{P}(\hat{i}\rightarrow j)||_2 = 
    %         ||L_i \cdot R_j-L_{\hat{i}} \cdot R_j||_2\\
    %         =||(L_i -L_{\hat{i}}) \cdot R_j||_2
    %     \end{gathered}
    % \end{equation}
    % If we look at the expectation of this, we see
    % \begin{equation}
    %     \begin{gathered}
    %         %\centering
    %         \mathbb{E}_j(||(L_i -L_{\hat{i}}) \cdot R_j||_2) = ||(L_i -L_{\hat{i}}) \cdot \frac{1}{k}\sum_{j\ne i}R_j||_2
    %     \end{gathered}
    % \end{equation}
    % Where $k$ is the number of neighbours we chose for the neural network input. Because the distance between points is described by the euclidean distance, we use this as our loss metric.

    % Symbolic Regression Explaination
    \subsection{Symbolic Regression}
        With our trained NODE, we have a black box that approximates the system that governs the movement of the target node in the embedded space. To make this more interpretable we then look to approximate this black box into a combination of known functions using symbolic regression\cite{pysr}.

        Symbolic regression is a type of regression analysis that searches a space of equations to find an expression that matches data. We use the genetic programming algorithm found in \cite{pysr}. The process for which can be seen in \autoref{symbreg}.

        \begin{figure}
            \centering
            \resizebox{\width}{!}{\input{SynthDataDiagrams/symbreg.tex}}
            \caption{Illustration of the genetic programming symbolic regression process.}
            \label{symbreg}
        \end{figure}

        The fitness of a candidate function is determined by comparing the predicted value to the given data, as well as the derivatives of the candidate equations to numerically found partial derivatives of the data variable.

        For this process we need to define a set of functions which the algorithm will use to generate a tree of functions to find the best approximation to our NNDE with. When using this method we are usually required to make two assumptions: that we have paired observations of $y(t), \frac{dy}{dt}$, and that the tree of expressions we will need is shallow. We will train the symbolic regression on the predictions of the neural network. By doing this, we can remove these assumptions\cite{kidger2022neural}.

    \subsection{Reconstructing the Temporal Network}
        To recover a final prediction, the point vector of the target node with the matrix $R'$, this gives a vector of probabilities of connections between the target node and the rest of the network. As the results can be interpreted as a probability, the most likely graph is given by setting any entry with a value of $0.5$ or higher to one (predict that the edge exists), and setting any entry with a value of less than $0.5$ to zero (predict that an edge does not exist).
\end{section}

\begin{section}{Data}
    In this paper we simulate three temporal sequences of undirected graphs. For each sequence we will be modelling the node in red, we will refer to this as the target node. For each sequence we train the model on the first 25 time steps and compare the predictions for the next 15 time steps. For clarity, each diagram \autoref{2community train series} \autoref{longTail train series} \autoref{3community train series} has been simplified and illustrates the movement of the target node in each sequence.

    \subsection{Two Community System}
        The 2 community network has communities of size 40 and 50, each of which are fully connected. The target node is initially fully connected to the 50 node community. In our simplified diagram \autoref{2community train series, a} the 50 node community is represented by the 5 fully connected nodes, while the 40 node community is represented by the 4 node component. At each time step one edge is removed between the target node and the larger community and replaced with an edge between the target node and the smaller community \autoref{2community train series, b}. This process is then continued \autoref{2community train series, c} for thirty-five time steps to generate our training and test data. Once we have this data in the form of a sequence of adjacency matrices, we find the singular value decomposition using the framework described in \autoref{svd}.

        Figures obtained from \cite{Emma2023}.
        \begin{figure}[H]
            \centering
            \centering
            \begin{subfigure}[c]{0.3\textwidth}
                \centering
                \resizebox{.6\width}{!}{\input{SynthDataDiagrams/2community1.tex}}
                \caption{$t=1$}
                \label{2community train series, a}
            \end{subfigure}
            \hfill
            \centering
            \begin{subfigure}[c]{0.3\textwidth}
                \centering
                \resizebox{.6\width}{!}{\input{SynthDataDiagrams/2community2.tex}}
                \caption{$t=2$}
                \label{2community train series, b}
            \end{subfigure}
            \hfill
            \centering
            \begin{subfigure}[c]{0.3\textwidth}
                \centering
                \resizebox{.6\width}{!}{\input{SynthDataDiagrams/2community3.tex}}
                \caption{$t=3$}
                \label{2community train series, c}
            \end{subfigure}
            \caption{First 3 time steps of the simplified, synthetic 2 community network. The red node represents the target node that we are aiming to model. The 4 node and 5 node fully connected components represent the 40 and 50 node communities of the sequence respectively. At each time step, an edge from the target node to the larger community is replaced with and edge from the target node to the smaller community.}
            \label{2community train series}
        \end{figure}

>>>>

% \begin{section}{Introduction}
%     Temporal networks are networks whose nodes and edges change in time. In this thesis we look to model temporal networks. That is, describe the changing temporal network using mathematical language. 
%     % Why temp nets are important
%     The modelling of temporal networks is an important task in many real world applications including symptom interactions for mental health \cite{jordan2020current,contreras2020temporal}, epidemiology \cite{masuda2013predicting}, and protein interactions \cite{lucas2021inferring,jin2009identifying}.

%     % Temp nets are dynam systs
%     Temporal networks can be seen as a system in which points, in our case nodes in a network, whose states, the edges connecting them, vary dependent in time, a dynamical system.

%     % Hard to find dynam systs because of discrete events 
%     Discovering the underlying equations governing these dynamical systems proves challenging. That is because changes in network structure are typically observed in the form of discrete jumps from one state to another; for example an edge between two nodes not being observed at the first time step then being observed at the next.

%     % We are proposing a solution
%     In this thesis, we propose a hybrid statistical and deep learning framework. That is a framework in which a neural network is used to model probabilistic values of the temporal network. This framework allows us to model temporal networks as continuous-time dynamical systems, discover a fitting set of differential equations describing it, and, exploiting that discovery, predict the time evolution of a temporal network.

%     % DEs are useful for modelling interacting states
%     Differential equations are useful for modelling systems where the state of one variable can effect the trajectories of other variables. We observe this behavior in temporal networks; nodes' connections within the network can influence the observation of edges between other nodes, for example the phenomenon observed in \cite{newman2001clustering,capocci2006preferential}, where a node is more likely to gain the connections the more connections it has (preferential attachment). With this in mind we might wish to draw on the rich mathematical literature of differential equation modelling.

%     % Temp net events are discrete
%     In the common representation of networks as binary-valued adjacency matrices, matrices where the observation or lack thereof, of an edge is represented as a 1 or 0 respectively, the events recorded in a temporal sequence of networks correspond to the appearance or the disappearance of link.

%     % Difficult to model discrete events
%     Because of the discrete nature of events, directly modelling the temporal networks as dynamical systems would require us to handle discrete jumps. The discontinuous character of their temporal evolution, make it challenging to use differential equations techniques.

%     % We propose a solution
%     Here, we overcome the discreteness problem by interpreting temporal networks as a well established statistical model for complex networks, which embeds nodes in a continuous, low-dimensional metric space by using a truncated singular value decomposition (Random Dot Product Graphs\cite{athreya2017statistical}). In this way we translate the hard problem of modelling discrete events in the space of networks to the easier problem of modelling continuous change in the embedding space.

%     % Our solution is interpretable because we use symb reg
%     We then define and use Neural Network Differential Equations (NNDE)\cite{SciML_C_Rak} to approximate the time evolution of the embedding space, and symbolic regression techniques to discover the functional form of the fitted NNDEs. These functional forms are interpretable (as they read as classic differential equations) and allow us to predict forward in time the evolution of the temporal networks.

%     % We propose a solution
%     In this manuscript, we show that the temporal network prediction problem can be successfully re-interpreted as a dynamical system modelling problem by taking the singular value decomposition of a sequence of adjacency matrices and training a NNDE to model an approximation to the underlying differential equation. We then go on to create a functional equation of this approximation.

%     % We will do some small tests on the proposed framework
%     We apply our proposed framework to three small example temporal networks with the hope of exploring the limitations and strengths of the proposed framework.

%     % The framework can and should be tweaked to make it optimised
%     The framework we are introducing is extremely flexible, and our research regarding the optimal structure of the Neural Networks used for the NNDEs is just started.
%     We are confident that future research can identify more fitting Neural Network structures than the simple one adopted here.
%     For this reason, we did not yet attempt to benchmark our model against other classic temporal network prediction methods.

%     % The framework can be applied to lots of fields
%     As it is completely general, we believe that the framework we are introducing can be usefully applied to areas of medicine, especially protein interaction networks; population dynamics for network ecology; and social network modelling. In particular, we discuss how specific domain knowledge relative to the prediction scenario can be taken into account, moving from NNDEs to Universal Differential Equations (UDEs).
% \end{section}

% \begin{section}{Literature Review}

%     \subsection{Temporal Networks}
%         % Use of temp nets
%         \subsubsection{Definitions}
%         Temporal networks are used to describe a series of interactions across time.
%         A temporal network $G$ is defined as:  
%         \begin{align}
%             G=\{ {_t} G=({_t} N, {_t} E)\} _{t \in 1\ldots T } 
%         \end{align}
%         $_t x \in {_t} N, {_t} e\in {_t} E$, where ${_t} x, {_t} e$ are nodes and edges at time $t$ respectively, and ${_t} N, {_t} E$ are the sets of nodes and edges at time $t$ respectively.
%         An example of this can be found in \autoref{method nets}.

%         %representations of temp nets as adj lists
%         This can be represented in many ways, for example as a series of interactions. The series of interactions is generally represented as a list of tuples containing the nodes which interacted and the time that the interaction took place. 
%         \begin{align}
%             G=\{(i, j, t)|\space i,j\in {_t} N, (i\leftrightarrow j)\in {_t}E \}
%         \end{align}

%         % representation of temp nets as adj mats
%         A temporal network can also be represented as a sequence of adjacency matrices. In the representation of a series of adjacency matrices, the columns and rows of the matrix each refer to each of the nodes. In this way, if an edge exists between nodes $i$ and $j$, the entry $(i,j)$ of the adjacency matrix will be 1; if an edge does not exist the entry will be 0.
%         \begin{align}
%             _t G &= {_t} A\\
%             _t A_{ij} &=  
%             \left\{
%                 \begin{array}{ll} 
%                 1 & (i\leftrightarrow j)\in {_t} E\\
%                 0 & otherwise
%                 \end{array}
%             \right.
%         \end{align}
%         The row $_t A_{i\cdot}$ describes the outgoing edges of the node $i$, while the column $_t A_{\cdot j}$ describes the incoming edges of the node $j$.

%         \subsubsection{Applications}
%         Temporal networks are present in many areas of research interest, ranging from symptom interactions in mental health\cite{jordan2020current,contreras2020temporal}, to epidemiology\cite{masuda2013predicting}, protein interactions\cite{lucas2021inferring,jin2009identifying}, and social networks\cite{moinet2015burstiness,hanneke2010discrete}.

%         % Overview of modelling attempts
%         Generally attempts to model temporal networks either model the change in node state or the change in network structure.

%         The node state describes some information about that node:
%         \begin{align}
%             _t x &= (m_1,\ldots, m_p)
%         \end{align}
%         Where $m_i$ is some metadata about the node $_t x$.

%         An example of modelling the change in a node state could be the severity of a symptom at a given time, as in \Citeauthor{contreras2020temporal}\cite{contreras2020temporal}. Where the node $_t x$ represents a given symptom and $m_1$ would represent the severity of that symptom.

%         % Intro contreras as attempt to model changes in nodes
%         \Citeauthor{contreras2020temporal}\cite{contreras2020temporal} model the change in node state over time while the structure of the network remains constant. The authors look at the severity of symptoms over time, i.e. the state of the nodes, where the nodes exist in a statically structured temporal network.

%         % Contreras methods
%         \Citeauthor{contreras2020temporal} use multilevel vector autoregression, a method that finds correlations between the current state of a variable and the states of variables at previous time steps \cite{singer2003applied}, to model how the current symptoms of people with paranoia might predict their future symptoms. These models were then used to create three networks that linked symptoms.

%         % Focal point of the paper is their temporal network
%         Of particular interest to this thesis is the temporal network created by having a fully connected network of all symptoms including self-loops, where a symptom connects to itself; this self loop comes about from a symptom being correlated with itself from one time step to the next. The edge weights represent the extent to which the severity of each symptom at time $(t)$ predicts the severity of itself and other symptoms at time $(t+1)$.

%         % Why this is the focal point
%         This framework keeps the overall structure of the network fixed throughout time and so is not particularly flexible, but is extremely human interpretable and can be used to extrapolate and predict into the future.

%         % intro to paper as attempt to model node state
%         In \cite{KARIMI20133476}\cite{KARIMI20133476}, the \Citeauthor{KARIMI20133476} look to model the change in node state while the network structure changes.

%         % How network is represented
%         \Citeauthor{KARIMI20133476} represent their temporal network as a series of interactions (a list of tuples in the form $(i,j,t)$, where $i,j$ are the nodes interacting and $t$ is the time the at which the interaction occurred).

%         % what state the paper attempts to model
%         The authors use a model to predict when a node will change from state 0 to state 1, in practice this could represent whether a user of a social network believes a rumour, or whether an ecological patch has been colonised.

%         % Model limitation
%         Notably, in this model, the node never changes back; this is done to keep the model analytically tractable, but does limit its usefulness in some contexts. In this model an organism could not become locally extinct in an ecological patch for example.

%         % Methods to model node state
%         To model the change in a nodes' state, the model uses a sliding time window and if the fraction of interactions with nodes of state 1 within that time window exceeds a given threshold, the node switches to state 1 as well.

%         % Limitation of model
%         In this work we aim to present a model for predicting the state of a node in the future. Instead, \Citeauthor{KARIMI20133476} aim to present a tool for understanding the spread of ideas, viruses, etc. within a network.

%         Other efforts focus on modelling the changing structure of the network as in \Citeauthor{sanna2021link}\cite{sanna2021link}, which looks to model $_t G$ as a whole.

%         % temp net representation
%         In this case the \Citeauthor{sanna2021link} represent the temporal network as a series of adjacency matrices, each matrix representing an observation of the changing network.

%         % Methods for embedding
%         The authors look to model the structure of edges by using a spectral embedding (SVD). A spectral embedding uses the eigenvectors of a network's adjacency matrix to find vectors that represent the nodes of the network. The spectral embedding transforms their temporal sequence into a sequence of observations of a latent space.

%         % What is a latent space
%         A latent space is a continuous space in which similar nodes (nodes with many of the same neighbours) are positioned near each other. In this space nodes are represented as points in a continuous, low dimensional space. Another feature of this latent space, is that the dot product of any two points in the space gives us the probability that the nodes related to those points are connected, a Random Dot Product Graph (RDPG)\cite{athreya2017statistical}.

%         % methods for time series
%         With this series of embeddings, the authors employ a variety of time series techniques to predict the network structure at future time steps.

%     \subsection{UDEs and NNDEs}
%         % Justification for DEs being useful
%         In some temporal networks the state of a node might influence the evolution of its neighbours. For example, in social networks we often observe that the more neighbours a node has, the more likely they are to gain more neighbours, a power law distribution of edges\cite{zhao2012multi,garg2009evolution}. We also see that the state of symptoms (occurrence, severity, and distress), can influence the states of other symptoms in patients undergoing chemotherapy \cite{papachristou2019network,kalantari2022network}, and for mental health disorders \cite{contreras2020temporal}. As well as in ecological networks where the interactions of species are often represented as a network, and the populations of one species may influence the population growth of another \cite{elton2001animal,volterra1927variazioni}; that is, the state (in this case the population) of a species (node), may influence the movement of another species.

%         % DEs have lots of research
%         With the observation that nodes in real world temporal networks may influence others, differential equations or difference equations seem appropriate for modelling the evolution of nodes; there is a rich body of literature exploring these methods use in modelling interacting variables.

%         % The data we have is discrete, which is a problem for DEs
%         One of the issues that the differential equation method encounters is the prevalence of discrete jumps of edges being observed and then not being observed. Because of this, we may consider exploring the possibility of using difference equations.

%         % time is continuous so good to have our model be cont too
%         We note that the progression of the network is generally continuous even if the events happen as discrete jumps. We may wish to preserve this continuous temporal progression in our model.

%         %This approach may overcome the problem of discrete observations, by defining a progression function that will only predict discrete formation or decay of edges\cite{hanneke2010discrete}. However, this imposes a significant limitation on the temporal granularity with which we can predict the state of the network. We would be limited to the granularity our data is collected at. This may be a problem if for example we had a series of weekly or monthly observations of networks we would not be able to predict interactions at each day. We can overcome both this limitation and the discrete jump limitation by using random dot product graphs. Interpolation may be possible in this case, but we would have no way of knowing if the interpolation had any basis in the real world.

%         % passino has discrete data but uses a method to view it as cont
%         Looking at the work of \Citeauthor{sanna2021link}\cite{sanna2021link}, the authors use RDPGs to approximate a temporal network as a matrix of probabilities of an edge existing between two nodes. In contrast to edges, these probabilities can evolve continuously, and so we can use differential equations to generate probabilistic networks at any temporal resolution we require. That is, we will be able to model the changes in the probabilities of edges continuously and from this, generate a network at any time step.

%         %  we can use the same method we can view data as dynam sys and use DEs
%         Using the method in \Citeauthor{sanna2021link}\cite{sanna2021link}, we can treat the sequence of embeddings as a dynamical system and then use differential equation modelling in a continuous space. Once we have found a latent space for each network, we know that each point in this space refers to a specific node in the network, and so, as the network changes, so too does the position of each of the points in the latent space. We then look to model this movement using differential equations.

%         % Can use a new NN structure to make this framework very general
%         Our aim for this framework to be as general as possible and note that there may not yet be enough domain knowledge to perform traditional differential equation modelling to the desired accuracy. For example in network ecology models can be useful, but not perfect. Hence, we might wish to build on these models rather than trying to create something entirely new. This can be achieved by using this partial domain knowledge using it to create a UDE\cite{SciML_C_Rak}.

%         \subsubsection{Theory}

%             % Overview of why UDEs are useful
%             Developed by \Citeauthor{SciML_C_Rak}\cite{SciML_C_Rak}, UDEs are a novel neural network architecture that aims to take the best of both traditional neural differential equation modelling and of flexible machine learning approaches to modelling.

%             If we consider a comparison to an Ordinary Differential Equation (ODE), the ODE is written as:
%             \begin{align}
%                 y(0) &= y_0 \\
%                 \frac{dy}{dt}(t)&=f(t,y(t))
%             \end{align}

%             The UDE will incorporate some component of $f(t,y(t))\approx \hat f(t,y(t))$, and will learn the difference with a Neural Network (NN) such that:
%             \begin{align}
%                 \label{UDE}
%                 y(0) &= y_0 \\
%                 \frac{dy}{dt}(t)&=\hat f(t,y(t))+NN(t,y(t), \theta)
%             \end{align}
%             Where $\theta$ is a vector of the parameters of the NN.
%             % Structure of UDE
%             UDEs combine domain knowledge in the form of a differential equation, with a NN. This hybrid machine learning model is then trained on the observed data.

%             % Overview of why UDEs are useful
%             Combining the two allows for most of the movement in the data to be captured by the differential equation, which contains the domain knowledge. In this way, the NN will only need to learn a theoretically simpler equation, given that a part of the system has already been captured.

%             % Why this structure is useful
%             This then allows the NN to be smaller, require fewer data, and be trained much faster than traditional NN models. Given the flexibility of NNs, the use of UDEs is natural when modelling a process that is not entirely understood\cite{kidger2022neural}.

%             Note that we can construct a special case of a UDE, an NNDE when $\hat f(t,y(t))=0$. This structure can be used when do not have any domain knowledge we wish to give the model.

%             % Software used for coding
%             There is an ecosystem of packages specifically designed for performing machine learning with UDEs. The ecosystem is called Scientific Machine Learning (SciML)\cite{SciML_C_Rak}. Largely, the SciML ecosystem has been optimised for flexibility and efficiency with respect to models available and training performance. Given there does not seem to be any other ecosystem with this feature, it seemed like 
%             an obvious choice.

%         \subsubsection{Applications}
%             Whilst there has been much interest in implementing UDEs and NNDEs, much of this work has been focussed on physics informed neural network and physics informed neural ordinary differential equations (PINNs and PINODEs) \cite{karniadakis2021physics,GAO2021110079,krishnapriyan2021characterizing,roehrl2020modeling}, and on improving modelling of fluids \cite{mahmoudabadbozchelou2021data,nguyen2022physics}. Alongside this, UDEs have gained popularity in the last few years, and there have been a number of studies exploring their usefulness in modeling the effect of restrictions due to COVID-19 on the virus' spread \cite{Dandekar2020.04.03.20052084}. Although there has been a large amount of research into the usefulness of these types of models, to the best of this author's knowledge they have never been applied to the problem of predicting temporal networks. This seems to be a natural fit for UDEs and NNDEs, given we may know relatively little about the processes that govern the evolution of temporal networks, and any information we do have can be incorporated into the model.

%     \subsection{Symbolic Regression}
%         % What is symbolic regression
%         Symbolic regression is a type of analysis that looks to take input variables and find a relationship between them that describes the output variables. For symbolic regression, we construct a set of mathematical operations that may be used to find this relationship. The algorithm then generates combinations of these equations to find an expression that matches the output variables. Because we construct the set of potential operations, the resulting expression is interpretable by humans (a functional equation), as opposed to a NN for example (a numeric equation).

%         % Why is symbolic regression useful
%         Having interpretable models that capture underlying relations is relevant to all fields of study, and so the applications of symbolic regression are incredibly varied.

%         % Where has symbolic regression been used 
%         Various methods of symbolic regression have been use in manufacturing systems, chemical systems, and tumor research\cite{can2011comparison,keith2021combining,yoshihara2013inferring}. As well research into improvements is still on going.

%         % How is this done
%         Finding a suitable functional expression can be achieved in many ways and remains active area of research. One of the most popular algorithms is genetic programming\cite{schmidt2009distilling}.

%         % How genetic programming works
%         Genetic programming involves constructing a search tree using an evolutionary algorithm to yield candidate equations. An evolutionary algorithm is an iterative process. It begins by randomly generating a set of candidate equations with operations from the given set of operations:
%         \begin{align}
%             Q_1=\{f_1,\dots,f_n\}
%         \end{align}
%         Each candidate function $f_i$ is tested and given a fitness score based on how well it performed. The higher a candidates' fitness, the more likely it is to be selected for the next iteration.

%         Suppose $f_1,f_2$ were selected for the next iteration. The evolutionary algorithm would then randomly generate new functions based on $f_1,f_2$ again with the given set of operations.
%         \begin{align}
%             Q_2=\{f_{1,1},\dots,f_{1,n}, f_{2,1},\dots,f_{2,n}\}
%         \end{align}
%         This process is repeated until a suitable functional equation is found.

%         For the genetic programming algorithm, the fitness of a candidate function is determined by how closely it matches the data as well as numerically calculated partial derivatives of each of the input variables.

%         % Why are the derivatives important
%         Comparing with these derivatives is useful as it helps to ensure that the produced function equations have some grounding in the real world process that generated the data. The search tree is then pruned by comparing the partial derivatives of the candidate equation, with partial derivatives calculated from data. This process is then iterated many times to generate a suitable equation.

%         % What symbolic regression aims to solve
%         The problem that symbolic regression aims to solve is to generate interpretable and ``meaningful'' equations from observed data.
